Consistent ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-08-22 manuscript/usses_water_ms.Rmd
tiny edits to methods in ms

## Data Analysis
Our main goal was to test whether the relationship between ANPP and soil moisture differed among the drought, control, and irrigation treatments.
Based on our own observations and previous work at our study site [@Blaisdell1958; @Dalgleish2011; @Adler2012], we chose to use cumulative volumetric water content from March through June as our metric of soil moisture (hereafter referred to simply as `soil moisture').
To achieve this goal, we fit a hierarchical regression model with log(ANPP) as the response variable and soil moisture as the sole predictor.
We log-transformed ANPP to account for heteroscedasticity.
Both log(ANPP) and soil moisture were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].

@@

\noindent{}where $\mu_{i}$ is the deterministic prediction from the regression model for observation *i*, which is associated with plot *j* and treatment year *t*. $\boldsymbol{\beta}$ is the vector of coefficients for the fixed effects in the design matrix $\textbf{X}$. Each row of the design matrix represents a single observation ($\textbf{x}_i$) and is a vector with the following elements: 1 for the intercept, a binary 0 or 1 if the treatment is "drought", a binary 0 or 1 if the treatment is "irrigation", the scaled value of volumetric water content (VWC), binary "drought" value times VWC, and binary "irrigation" value times VWC. Thus, our model treats "control" observations as the main treatment and then estimates intercept and slope offsets for the "drought" and "irrigation" treatments. In reference to our model, the hypotheses we wish to test are: (1) the coefficient for drought$\times$VWC is positive and different from zero, and (2) the coefficient for irrigation$\times$VWC is negative and different from zero

To account for the fact that observations within plots and years are not independent, we include three random effects. Specifically, we include plot-specific offsets ($\boldsymbol{\gamma}$) for the intercept and slope terms and year-specific intercept offsets ($\eta_t$). The covariate vector $\textbf{z}_i$ for each observation *i* has two elements: a 1 for the intercept and the scaled value of VWC for that plot and year. The plot-specific coefficients are modeled hierarchically, where plot level coefficients are drawn from a multivariate normal distribution with mean 0 and a variance-covariance strucutre that allows the intercept and slope terms to be correlated:

@@

```{r get_results, echo=FALSE}
betas <- read.csv("../results/beta_summaries.csv")
```

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3), but the effect is associated with high uncertainty, with credible intervals that overlap zero, regardless of treatment (Table 1).
Ecosystem functional response was similar among treatments (Table 1; Fig. 3), but there is evidence that the slope for the irrigation treatment is less than the slope for the control treatment.
This evidence comes from interpreting the posterior distribution of the slope offset for the irrigation treatment, from which we calculate a 98% probability that the estimate is less than zero (Fig. 3A, right panel).
There was no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. A2-2).

>>>>

## Data Analysis
Our main goal was to test whether the relationship between ANPP and soil moisture differed among the drought, control, and irrigation treatments.
Based on our own observations and previous work at our study site [@Blaisdell1958; @Dalgleish2011; @Adler2012], we chose to use cumulative volumetric water content from March through June as our metric of soil moisture (hereafter referred to as `VWC').
To achieve this goal, we fit a generalized linear mixed effects regression model with log(ANPP) as the response variable and VWC and treatment as fixed effects.
Plot and year of treatment were included as random effects to account for non-independence of observations, as described below.
We log-transformed ANPP to account for heteroscedasticity.
Both log(ANPP) and VWC were standardized to have mean 0 and unit variance before fitting the model [i.e., $(x_i - \bar{x})/\sigma_x$].

@@

\noindent{}where $\mu_{i}$ is the deterministic prediction from the regression model for observation *i*, which is associated with plot *j* and treatment year *t*. $\boldsymbol{\beta}$ is the vector of coefficients for the fixed effects in the design matrix $\textbf{X}$. Each row of the design matrix represents a single observation ($\textbf{x}_i$) and is a vector with the following elements: 1 for the intercept, a binary 0 or 1 if the treatment is "drought", a binary 0 or 1 if the treatment is "irrigation", the scaled value of VWC, binary "drought" value times VWC, and binary "irrigation" value times VWC. Thus, our model treats "control" observations as the main treatment and then estimates intercept and slope offsets for the "drought" and "irrigation" treatments. In reference to our model, the hypotheses we wish to test are:

\begin{description}
\item [H1] The coefficient for drought$\times$VWC is positive and different from zero.
\item [H2] The coefficient for irrigation$\times$VWC is negative and different from zero.
\end{description}

To account for the fact that observations within plots and years are not independent, we include two random effects. Specifically, we include plot-specific offsets ($\boldsymbol{\gamma}$) for the intercept and slope terms and year-specific intercept offsets ($\eta_t$). The covariate vector $\textbf{z}_i$ for each observation *i* has two elements: a 1 for the intercept and the scaled value of VWC for that plot and year. The plot-specific coefficients are modeled hierarchically, where plot level coefficients are drawn from a multivariate normal distribution with mean 0 and a variance-covariance strucutre that allows the intercept and slope terms to be correlated:

@@

<!---
```{r get_results, echo=FALSE}
betas <- read.csv("../results/beta_summaries.csv")
```
--->

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP (Table 1; Fig. 3), but the effect is associated with high uncertainty, with credible intervals that overlap zero, regardless of treatment (Table 1).
Ecosystem functional response was similar among treatments (Table 1; Fig. 3B), but there is evidence that the slope for the irrigation treatment is less than the slope for the control treatment.
This evidence comes from interpreting the posterior distribution of the slope offset for the irrigation treatment, from which we calculate a 98% probability that the estimate is less than zero (Fig. 3A, right panel).
There was no evidence that the treatment effects became more important over time because there was no directional trend in the random year effects (Fig. A2-2).

### 2017-08-22 manuscript/usses_water_ms.tex
tiny edits to methods in ms

Our main goal was to test whether the relationship between ANPP and soil
moisture differed among the drought, control, and irrigation treatments.
Based on our own observations and previous work at our study site
\citep{Blaisdell1958, Dalgleish2011, Adler2012}, we chose to use
cumulative volumetric water content from March through June as our
metric of soil moisture (hereafter referred to simply as `soil
moisture'). To achieve this goal, we fit a hierarchical regression model
with log(ANPP) as the response variable and soil moisture as the sole
predictor. We log-transformed ANPP to account for heteroscedasticity.
Both log(ANPP) and soil moisture were standardized to have mean 0 and
unit variance before fitting the model {[}i.e.,
\((x_i - \bar{x})/\sigma_x\){]}.

@@

\noindent{}where \(\mu_{i}\) is the deterministic prediction from the
regression model for observation \emph{i}, which is associated with plot
\emph{j} and treatment year \emph{t}. \(\boldsymbol{\beta}\) is the
vector of coefficients for the fixed effects in the design matrix
\(\textbf{X}\). Each row of the design matrix represents a single
observation (\(\textbf{x}_i\)) and is a vector with the following
elements: 1 for the intercept, a binary 0 or 1 if the treatment is
``drought'', a binary 0 or 1 if the treatment is ``irrigation'', the
scaled value of volumetric water content (VWC), binary ``drought'' value
times VWC, and binary ``irrigation'' value times VWC. Thus, our model
treats ``control'' observations as the main treatment and then estimates
intercept and slope offsets for the ``drought'' and ``irrigation''
treatments. In reference to our model, the hypotheses we wish to test
are: (1) the coefficient for drought\(\times\)VWC is positive and
different from zero, and (2) the coefficient for irrigation\(\times\)VWC
is negative and different from zero

To account for the fact that observations within plots and years are not
independent, we include three random effects. Specifically, we include
plot-specific offsets (\(\boldsymbol{\gamma}\)) for the intercept and
slope terms and year-specific intercept offsets (\(\eta_t\)). The
covariate vector \(\textbf{z}_i\) for each observation \emph{i} has two
elements: a 1 for the intercept and the scaled value of VWC for that
plot and year. The plot-specific coefficients are modeled
hierarchically, where plot level coefficients are drawn from a
multivariate normal distribution with mean 0 and a variance-covariance
strucutre that allows the intercept and slope terms to be correlated:

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP
(Table 1; Fig. 3), but the effect is associated with high uncertainty,
with credible intervals that overlap zero (Table 1). Ecosystem
functional response was similar among treatments (Table 1; Fig. 3), but
there is evidence that the slope for the irrigation treatment is less
than the slope for the control treatment. This evidence comes from
interpreting the posterior distribution of the slope offset for the
irrigation treatment, from which we calculate a 98\% probability that
the estimate is less than zero (Fig. 3A, right panel). There was no
evidence that the treatment effects became more important over time
because there was no directional trend in the random year effects (Fig.
A2-2).

@@

The results of our five year experiment in a sagebrush steppe conform to
two of our four predictions, depending on treatment. Under chronic
drought, neither ecosystem functional response nor community composition
changed (Fig. 1, top left). But under irrigation, ecosystem functional
response was different from the control treatment while community
composition remained unchanged (Fig. 1, top right). Thus, we found some
evidence for a saturating response at high water availability (Fig. 3),
but the response of ANPP to soil moisture was weak overall (Table 1).

>>>>

Our main goal was to test whether the relationship between ANPP and soil
moisture differed among the drought, control, and irrigation treatments.
Based on our own observations and previous work at our study site
\citep{Blaisdell1958, Dalgleish2011, Adler2012}, we chose to use
cumulative volumetric water content from March through June as our
metric of soil moisture (hereafter referred to as `VWC'). To achieve
this goal, we fit a generalized linear mixed effects regression model
with log(ANPP) as the response variable and VWC and treatment as fixed
effects. Plot and year of treatment were included as random effects to
account for non-independence of observations, as described below. We
log-transformed ANPP to account for heteroscedasticity. Both log(ANPP)
and VWC were standardized to have mean 0 and unit variance before
fitting the model {[}i.e., \((x_i - \bar{x})/\sigma_x\){]}.

@@

\noindent{}where \(\mu_{i}\) is the deterministic prediction from the
regression model for observation \emph{i}, which is associated with plot
\emph{j} and treatment year \emph{t}. \(\boldsymbol{\beta}\) is the
vector of coefficients for the fixed effects in the design matrix
\(\textbf{X}\). Each row of the design matrix represents a single
observation (\(\textbf{x}_i\)) and is a vector with the following
elements: 1 for the intercept, a binary 0 or 1 if the treatment is
``drought'', a binary 0 or 1 if the treatment is ``irrigation'', the
scaled value of VWC, binary ``drought'' value times VWC, and binary
``irrigation'' value times VWC. Thus, our model treats ``control''
observations as the main treatment and then estimates intercept and
slope offsets for the ``drought'' and ``irrigation'' treatments. In
reference to our model, the hypotheses we wish to test are:

\begin{description}
\item [H1] The coefficient for drought$\times$VWC is positive and different from zero.
\item [H2] The coefficient for irrigation$\times$VWC is negative and different from zero.
\end{description}

To account for the fact that observations within plots and years are not
independent, we include two random effects. Specifically, we include
plot-specific offsets (\(\boldsymbol{\gamma}\)) for the intercept and
slope terms and year-specific intercept offsets (\(\eta_t\)). The
covariate vector \(\textbf{z}_i\) for each observation \emph{i} has two
elements: a 1 for the intercept and the scaled value of VWC for that
plot and year. The plot-specific coefficients are modeled
hierarchically, where plot level coefficients are drawn from a
multivariate normal distribution with mean 0 and a variance-covariance
strucutre that allows the intercept and slope terms to be correlated:

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP
(Table 1; Fig. 3), but the effect is associated with high uncertainty,
with credible intervals that overlap zero, regardless of treatment
(Table 1). Ecosystem functional response was similar among treatments
(Table 1; Fig. 3B), but there is evidence that the slope for the
irrigation treatment is less than the slope for the control treatment.
This evidence comes from interpreting the posterior distribution of the
slope offset for the irrigation treatment, from which we calculate a
98\% probability that the estimate is less than zero (Fig. 3A, right
panel). There was no evidence that the treatment effects became more
important over time because there was no directional trend in the random
year effects (Fig. A2-2).

@@

The results of our five year experiment in a sagebrush steppe conform to
two of our four predictions, depending on treatment. Under chronic
drought, neither ecosystem functional response nor community composition
changed (Fig. 1, top left). But under irrigation, ecosystem functional
response was different from the control treatment while community
composition remained unchanged (Fig. 1, top right). Thus, we found some
evidence for a saturating response at high water availability (Fig. 3),
but the response of ANPP to soil moisture was weak overall, regardless
of treatment (Table 1).
